Growth Curves Paper

### 2020-03-05 R Analysis Scripts/CFU Data.Rmd
moved here

#Data gathering functions:
GetGrowthCurveReplicate<- function (sample){
  bugCounts<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/Spatial paper Data/GrowthCurves/", sample, ".csv"),header = TRUE)
  bugCounts<- na.omit(bugCounts)
  bugCounts<-bugCounts %>% mutate(logCFU = log10(CFU))
  bugCounts$sample<-sample
  return(bugCounts)
}
GetGrowthCurveCondition<-function(datalist){
  mappedData<-map(datalist, GetGrowthCurveReplicate)#this line reads all the files on the list
  boundData<-bind_rows(mappedData)
  colnames(boundData)[5] <- "strain"
  colnames(boundData)[8] <- "condition"
  return(boundData)
}
#Graphing functions
##this one graphs the log values
GraphGrowthCurveLog<-function (rawData, statsData){
    ggplot()+
    geom_point(data=rawData, aes(x=time, y=logCFU,color = bug, shape = strain),   size=2)+
    #geom_point(data=statsData, aes(x=time, y=logCFUmean,  color = bug, shape = strain), size=2)+
    geom_line(data=statsData, aes(x=time, y=logCFUmean,  color = bug, linetype = strain), size=2)+
    #geom_errorbar(data=statsData, aes(x = time, ymin = logCFUmean-logCFUsd, ymax = logCFUmean+logCFUsd, color = bug, linetype = strain), width = 0, size = 1)+
    geom_errorbar(data=statsData, aes(x = time, ymin = logCFUmean-logCFUse, ymax = logCFUmean+logCFUse, color = bug, linetype = strain), width = 0, size = 1)+
    labs(title =paste("Growth in SCFM2" ),  x = "Time (hours)",shape = "Strain",linetype = "Strain", color = "Bug")+
    facet_grid(.~condition)
}
##this one graphs the normal values
GraphGrowthCurve<-function (rawData, statsData){
    ggplot()+ 
    geom_point(data=rawData, aes(x=time, y=CFU,color = bug, shape = strain),   size=2)+
    geom_line(data=statsData, aes(x=time, y=CFUmean,  color = bug, linetype = strain), size=2)+
    geom_errorbar(data=statsData, aes(x = time, ymin = CFUmean-CFUsd, ymax = CFUmean+CFUsd, color = bug, linetype = strain), width = 0, size = 1)+
    labs(title =paste("Growth in SCFM2" ),  x = "Time (hours)",shape = "strain", color = "Species")+
    scale_y_log10(name=expression(paste('CFUs per mL' )), breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    facet_grid(.~condition)
}

@@

```{r CFU results}
#these are the list with the spreadsheets of experiments under different conditions
PA14 = c('2-11-19','2-7-19','2-5-19')
#get data
PA14Data<-GetGrowthCurveCondition(PA14)
#do stats
PA14DataStats<-PA14Data%>% group_by(condition, bug, time,strain) %>% summarise(CFUmean = mean (CFU), CFUsd = sd(CFU), CFUse = std.error(CFU), logCFUmean = mean (logCFU), logCFUsd = sd(logCFU),logCFUse = std.error(logCFU))
#Graph logs
GraphGrowthCurveLog(PA14Data,PA14DataStats)
#Graph regular numbers
GraphGrowthCurve(PA14Data,PA14DataStats)

```{r Confocal vs CFUs data wrangling}
#first pull confocal data:
##functions to read data
MapResults<- function(listSamples, reps){
  blah <-deparse(substitute(listSamples))
  mappedData<-map2(listSamples, reps, GetResults)
  boundData<-bind_rows(mappedData)
  boundData<-boundData %>% mutate(Strain = blah)
  boundData<-gather(boundData,"Color","Value",-Slice,-Position,-Sample,-TimePoint,-Strain) %>% mutate(LogValue=ifelse(Value>0,log10(Value),0))
  boundData<-boundData %>% mutate(Condition = recode(Position, "0" = "Sa Mono", "1" = "Pa Mono", "2" = "Pa:Sa 1:1", "3" = "Pa:Sa 10:1", "4" = "Pa:Sa 1:100", "5" = "Pa:Sa 1:10" )) %>% mutate(Species = recode(Color, "Red" = "Sa", "Green" = "Pa")) %>% mutate(ZDistance = Slice *0.44)
  boundData$Condition<-as.factor(boundData$Condition)
  boundData$Condition<-factor(boundData$Condition, levels = c("Sa Mono", "Pa Mono", "Pa:Sa 1:1", "Pa:Sa 1:10", "Pa:Sa 1:100","Pa:Sa 10:1"))
  boundData$Species<-as.factor(boundData$Species)
  boundData$Strain<-as.factor(boundData$Strain)
  return(boundData)
  }
ReadResultsFile<-function(file, position, timepoint){
  datos <-read.csv(file, header = FALSE)
  colnames(datos)<-c('Red', 'Green', 'Slice')
  datos<-datos %>% mutate(Position = position)
  datos<-datos %>% mutate(TimePoint = timepoint)
  return(datos)
}
GetResults<-function(sample, samplename){
  folder<- paste0('/Users/jupabago/Documents/Whiteley/PROJECTS/Spatial paper Data/Microscopy Data/Other Data1/',sample,'/results')#the path needs to be "Other Data1"
  fileList<-list.files(path = folder, full.names = TRUE)
  growthDf<-data.frame()
  for (position in 0:5){
    currentPosition<-fileList[grep(paste0("p",position,"."),fileList)]
    for (timepoint in 0:17){
      currentTimePoint<-currentPosition[grep(paste0("t",GetNumber(timepoint),"_"),currentPosition)]
      growthDf<-rbind(growthDf,ReadResultsFile(currentTimePoint, position, timepoint))}}
  growthDf<-growthDf %>% mutate(Sample = samplename)
return(growthDf)
}

@@

##remove conditions I didn't do in the scope and fix strain names:
DataComparisonCFU<-PA14DataStats %>% filter (condition == "Static") %>% select(bug, strain, CFUmean, logCFUmean, condition, time)%>% as.data.frame() 
DataComparisonCFU<-DataComparisonCFU%>% mutate(Strain = recode(DataComparisonCFU$strain, wt = "PA14wt", mut = "PA14mut"))

@@

#Now we go to Sa
SaLinearModel <- lm(MeanLogTotals ~logCFUs, data = comparisonDf %>% filter (Species == "Sa"))
coefficients(SaLinearModel)
get_regression_summaries(SaLinearModel, print = TRUE)

SaLinearModel1 <- lm(MeanLogTotals ~logCFUs, data = comparisonDf %>% filter (Species == "Sa")%>% filter(!time == 0))
coefficients(SaLinearModel1)
get_regression_summaries(SaLinearModel1, print = TRUE)

ggplot()+
  geom_point(data=comparisonDf %>% filter(Species == "Sa"), aes(x=logCFUs,y=MeanLogTotals,  color = time, shape = Strain),size=4)+
  geom_abline(intercept =25.73, slope = -2.44, color = "red", size = 1, linetype = "dashed" )+
  scale_color_gradient(low="blue", high="red")+
  annotate("text", x = 7.9, y = 4.75, label = expression(paste("y=-2.44x+25.73  ", R^2, "= 0.619")), color = "red", size = 8)+
  labs(title =expression(paste(italic("S. aureus"), " CFU vs Confocal Biomass")), x = "Log CFUs",y = "Biomass from Scope")

>>>>

#Data gathering functions:
GetGrowthCurveReplicate<- function (sample){
  bugCounts<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/GrowthCurves/", sample, ".csv"),header = TRUE)
  bugCounts<- na.omit(bugCounts)
  bugCounts<-bugCounts %>% mutate(logCFU = log10(CFU))
  bugCounts$sample<-sample
  return(bugCounts)
}
GetGrowthCurveCondition<-function(datalist){
  mappedData<-map(datalist, GetGrowthCurveReplicate)#this line reads all the files on the list
  boundData<-bind_rows(mappedData)
  colnames(boundData)[5] <- "strain"
  colnames(boundData)[8] <- "condition"
  return(boundData)
}
#Graphing functions
##this one graphs the log values
GraphGrowthCurveLog<-function (rawData, statsData){
    staticPlot<-ggplot()+
      geom_line(data=statsData%>% filter(bug == "Sa"& condition == "Static"), aes(x=time, y=logCFUmean, linetype = strain), size=1)+
      geom_errorbar(data=statsData%>% filter(bug == "Sa"& condition == "Static"), aes(x = time, ymin = logCFUmean-logCFUse, ymax = logCFUmean+logCFUse), width = 0, size = 1)+
      scale_linetype_manual(values = c( "dotted","solid"), labels = c("- HQNO", "wild type"))+
      labs(x = "Time (hours)", y = "log CFU", linetype = "Pa strain")+
      scale_y_continuous(limits = c(4,9))+
      theme_cowplot(12)

    wellmixedPlot<-ggplot()+
      geom_line(data=statsData%>% filter(bug == "Sa" & condition == "Shaking"), aes(x=time, y=logCFUmean, linetype = strain), size=1)+
      geom_errorbar(data=statsData%>% filter(bug == "Sa"& condition == "Shaking"), aes(x = time, ymin = logCFUmean-logCFUse, ymax = logCFUmean+logCFUse), width = 0, size = 1)+
      scale_linetype_manual(values = c( "dotted", "solid"), labels = c("- HQNO","wild type"))+
      labs(x = "Time (hours)", y = "log CFU", linetype = "Pa strain")+
      scale_y_continuous(limits=c(4,9))+
      theme_cowplot(12)
    legend <- get_legend(staticPlot + theme(legend.box.margin = margin(0, 0, 0, 12)))
    prow <- plot_grid(wellmixedPlot+theme(legend.position="none"),staticPlot+theme(legend.position="none"),align='vh',labels = c("A", "B"),hjust = -1,nrow = 1)
    plot_grid(prow,legend,rel_widths = c(3, .4))
}
GraphGrowthCurveLog1<-function (rawData, statsData){
    wtPlot<-ggplot()+
      geom_line(data=statsData%>% filter(bug == "Sa"& strain == "wt"), aes(x=time, y=logCFUmean, linetype = condition), size=1)+
      geom_errorbar(data=statsData%>% filter(bug == "Sa"& strain == "wt"), aes(x = time, ymin = logCFUmean-logCFUse, ymax = logCFUmean+logCFUse), width = 0, size = 0.5)+
      scale_linetype_manual(values = c( "solid","dotted"), labels = c("well-mixed","static"))+
      labs(x = "Time (hours)", y = "log CFU", linetype = "Condition")+
      scale_y_continuous(limits = c(4,9.2))+
      scale_x_continuous(breaks = c(4,8,12,16))+
      theme_cowplot(12)

    hqnoPlot<-ggplot()+
      geom_line(data=statsData%>% filter(bug == "Sa" & strain == "mut"), aes(x=time, y=logCFUmean, linetype = condition), size=1)+
      geom_errorbar(data=statsData%>% filter(bug == "Sa"& strain == "mut"), aes(x = time, ymin = logCFUmean-logCFUse, ymax = logCFUmean+logCFUse), width = 0, size = 0.5)+
      scale_linetype_manual(values = c( "solid","dotted"), labels = c("well-mixed","static"))+
      labs(x = "Time (hours)", y = "log CFU", linetype = "Condition")+
      scale_y_continuous(limits=c(4,9.2))+
      scale_x_continuous(breaks = c(4,8,12,16))+
      theme_cowplot(12)
    legend <- get_legend(wtPlot + theme(legend.box.margin = margin(0, 0, 0, 12)))
    prow <- plot_grid(wtPlot+theme(legend.position="none"),hqnoPlot+theme(legend.position="none"),align='vh',labels = c("A", "B"),hjust = -1,nrow = 1)
    plot_grid(prow,legend,rel_widths = c(3, .4))
    }
##this one graphs the normal values
GraphGrowthCurve<-function (rawData, statsData){
    ggplot()+ 
    geom_point(data=rawData, aes(x=time, y=CFU,color = bug, shape = strain),   size=2)+
    geom_line(data=statsData, aes(x=time, y=CFUmean,  color = bug, linetype = strain), size=1)+
    geom_errorbar(data=statsData, aes(x = time, ymin = CFUmean-CFUsd, ymax = CFUmean+CFUsd, color = bug, linetype = strain), width = 0, size = 1)+
    labs(title =paste("Growth in SCFM2" ),  x = "Time (hours)",shape = "strain", color = "Species")+
    scale_y_log10(name=expression(paste('CFUs per mL' )), breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    facet_grid(.~condition)
}

@@

```{r CFU results}
#these are the list with the spreadsheets of experiments under different conditions
PA14 = c('2-11-19','2-7-19','2-5-19')
#get data
PA14Data<-GetGrowthCurveCondition(PA14)
PA14Data$strain <- factor(PA14Data$strain, levels = c("wt", "mut"))

#do stats
PA14DataStats<-PA14Data%>% group_by(condition, bug, time,strain) %>% summarise(CFUmean = mean (CFU), CFUsd = sd(CFU), CFUse = std.error(CFU), logCFUmean = mean (logCFU), logCFUsd = sd(logCFU),logCFUse = std.error(logCFU))
#Graph logs
GraphGrowthCurveLog(PA14Data,PA14DataStats)
GraphGrowthCurveLog1(PA14Data,PA14DataStats)
#Graph regular numbers
GraphGrowthCurve(PA14Data,PA14DataStats)
```

```{r Confocal vs CFUs data wrangling}
#first pull confocal data:
##functions to read data
MapResults<- function(listSamples, reps){
  blah <-deparse(substitute(listSamples))
  mappedData<-map2(listSamples, reps, GetResults)
  boundData<-bind_rows(mappedData)
  boundData<-boundData %>% mutate(Strain = blah)
  boundData<-gather(boundData,"Color","Value",-Slice,-Position,-Sample,-TimePoint,-Strain) %>% mutate(LogValue=ifelse(Value>0,log10(Value),0))
  boundData<-boundData %>% mutate(Condition = recode(Position, "0" = "Sa Mono", "1" = "Pa Mono", "2" = "Pa:Sa 1:1", "3" = "Pa:Sa 10:1", "4" = "Pa:Sa 1:100", "5" = "Pa:Sa 1:10" )) %>% mutate(Species = recode(Color, "Red" = "Sa", "Green" = "Pa")) %>% mutate(ZDistance = Slice *0.44)
  boundData$Condition<-as.factor(boundData$Condition)
  boundData$Condition<-factor(boundData$Condition, levels = c("Sa Mono", "Pa Mono", "Pa:Sa 1:1", "Pa:Sa 1:10", "Pa:Sa 1:100","Pa:Sa 10:1"))
  boundData$Species<-as.factor(boundData$Species)
  boundData$Strain<-as.factor(boundData$Strain)
  return(boundData)
  }
ReadResultsFile<-function(file, position, timepoint){
  datos <-read.csv(file, header = FALSE)
  colnames(datos)<-c('Red', 'Green', 'Slice')
  datos<-datos %>% mutate(Position = position)
  datos<-datos %>% mutate(TimePoint = timepoint)
  return(datos)
}
GetResults<-function(sample, samplename){
  folder<- paste0('/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/Microscopy Data/Other Data1/',sample,'/results')#the path needs to be "Other Data1"
  fileList<-list.files(path = folder, full.names = TRUE)
  growthDf<-data.frame()
  for (position in 0:5){
    currentPosition<-fileList[grep(paste0("p",position,"."),fileList)]
    for (timepoint in 0:17){
      currentTimePoint<-currentPosition[grep(paste0("t",GetNumber(timepoint),"_"),currentPosition)]
      growthDf<-rbind(growthDf,ReadResultsFile(currentTimePoint, position, timepoint))}}
  growthDf<-growthDf %>% mutate(Sample = samplename)
return(growthDf)
}

@@

##remove stirred condition I didn't do in the scope and fix strain names:
DataComparisonCFU<-PA14DataStats %>% filter (condition == "Static") %>% select(bug, strain, CFUmean, logCFUmean, condition, time)%>% as.data.frame() 
DataComparisonCFU<-DataComparisonCFU%>% mutate(Strain = recode(DataComparisonCFU$strain, wt = "PA14wt", mut = "PA14mut"))

@@

#Now we go to Sa

SaLinearModel <- lm(MeanLogTotals ~logCFUs, data = comparisonDf %>% filter (Species == "Sa")%>% filter(!time == 0))
coefficients(SaLinearModel)
get_regression_summaries(SaLinearModel, print = TRUE)

ggplot()+
  geom_point(data=comparisonDf %>% filter(Species == "Sa"), aes(x=logCFUs,y=MeanLogTotals,  color = time, shape = Strain),size=4)+
  geom_abline(intercept =25.73, slope = -2.44, color = "red", size = 1, linetype = "dashed" )+
  scale_color_gradient(low="blue", high="red")+
  annotate("text", x = 7.9, y = 4.5, label = expression(paste("y=-2.44x+25.73  ", R^2, "= 0.619")), color = "red", size = 8)+
  labs(title =expression(paste(italic("S. aureus"), " CFU vs Confocal Biomass")), x = "Log CFUs",y = "Biomass from CLSM")+
  theme_cowplot(12)
